\LARGE \bf Stochastic Uncertainty Propagation in Power System Dynamics using Measure-valued Proximal Recursions

### 2022-08-12 Papers/Stochastic_Uncertainty_Propagation_TPS/RebuttalRound2-TPWRS-01391-2021R1.tex

\noindent
{\bf R4.1. Comments by reviewer}:\\
{\em The authors have successfully addressed most of the reviewer's previous concerns. However, the reviewer still has some questions, mostly regarding the numerical simulations.}

{\nib \blue{Thank you. Please find our itemized responses to the questions related to numerical simulations.}}

\noindent
{\bf R4.2. Comments by reviewer}:\\
{\em It is stated that ``The Fig 9 shows that the computational time needed to perform the proximal updates, for all $t_k$, are indeed orders-of-magnitude faster than $h$."
This statement doesn't seem to hold, since $h=10^{-3}$ and the computational time lies within the range of around $10^{-3}$ to $10^{-2}$. How does the computational time needed to perform the proximal updates relate to time step $h$? If they are not strongly related, does that mean the step size $h$ is constrained to be greater than that computational time?}

{\nib \blue{The reviewer is right, we should have been more careful in phrasing that sentence. We meant to say that the computational time needed to perform a proximal update is \ul{statistically} comparable (in the big O sense) to $h$. The reason why the precise computational time is probabilistic is twofold: random sampling, and randomness in computational latency, power and memory footprint. Fig. 9 shows that $h=10^{-3}$ and the computational time needed to perform a single proximal update is $O(10^{-3})$ with high probability in the sense most time-steps incur $O(10^{-3})$ time. In Fig. 9, while instances when computational times are $O(10^{-2})$ do occur for Case I, they are rare. \ul{In the present revision, we have now rephrased the corresponding text (highlighted in blue) toward the end of Sec. V-A}.

The proximal updates involve certain block coordinate ascent recursions in the dual co-ordinates $\bm{y},\bm{z}$; see lines 13-21 in Algorithm 1. This recursion is guaranteed to be contractive (hence existence, uniqueness and convergence are guaranteed) in the product orthant cone: $(\bm{y},\bm{z})\in\mathbb{R}^{N}_{>0}\times \mathbb{R}^{N}_{>0}$, see ref. [41, Sec. III-B, III-C] and especially Corollary 4 in that reference. We have cited that reference in this manuscript's Sec. IV.}}

### 2022-08-12 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS_Rev2.tex
Rebuttal round 2

At each fixed physical time $t_{k}=kh$, $k\in\mathbb{N}$, where $h$ is the fixed step-size, we perform the proximal update $\varrho_{k-1}\mapsto\varrho_{k}$ via Algorithm 1. This proximal update has no analytical solution and instead requires execution of a contractive fixed point recursion. The physical time $t_{k}$ is ``frozen" (zero-order-hold) during the fixed point recursions underlying this one-step proximal update. For the proposed method to be practical, we therefore, need to demonstrate that the computational timescale is faster than the physical dynamics timescale, i.e., computational times needed to complete the zero-order-hold fixed point recursions are orders-of-magnitude faster than $h$, the step-size of physical time. On the other hand, theoretical consistency (see the last paragraph of Sec. \ref{subsec:ProxUpdate}) requires that $h$ itself should be small. Thus, even if one has a provably convergent algorithm that can perform the one step proximal update, it is not clear \emph{a priori} if the computational time needed to do so can be much faster than $h$. If not, then the method would be statistically consistent but practically not useful. The Fig \ref{fig:CompTimeIEEE14bus} shows that the computational time needed to perform the proximal updates, for all $t_{k}$, are indeed orders-of-magnitude faster than $h$.

>>>>

At each fixed physical time $t_{k}=kh$, $k\in\mathbb{N}$, where $h$ is the fixed step-size, we perform the proximal update $\varrho_{k-1}\mapsto\varrho_{k}$ via Algorithm 1. This proximal update has no analytical solution and instead requires execution of a contractive fixed point recursion. The physical time $t_{k}$ is ``frozen" (zero-order-hold) during the fixed point recursions underlying this one-step proximal update. For the proposed method to be practical, we therefore, need to demonstrate that {\blue{the computational timescale is at least statistically comparable (in the order of magnitude sense) to the physical dynamics timescale. In other words, the computational times needed to complete the zero-order-hold fixed point recursions are, with high probability, the same orders-of-magnitude as $h$, the step-size of physical time. The precise computational time is probabilistic, not just due to random sampling, but also due to randomness in power and memory footprint.}}

%is faster than the physical dynamics timescale, i.e., computational times needed to complete the zero-order-hold fixed point recursions are orders-of-magnitude faster than $h$, the step-size of physical time. On the other hand, theoretical consistency (see the last paragraph of Sec. \ref{subsec:ProxUpdate}) requires that $h$ itself should be small. Thus, even if one has a provably convergent algorithm that can perform the one step proximal update, it is not clear \emph{a priori} if the computational time needed to do so can be much faster than $h$. If not, then the method would be statistically consistent but practically not useful.

{\blue{In Fig \ref{fig:CompTimeIEEE14bus}, $h=10^{-3}$ and the computational time needed to perform a single proximal update, \emph{with high probability}, is $O(10^{-3})$ in the sense most time-steps incur $O(10^{-3})$ time. While instances when computational times are $O(10^{-2})$ do occur for Case I in Fig \ref{fig:CompTimeIEEE14bus}, but are rare.}}
